---
title: "SDS 315 - Homework 2"
author: "Brandon Duvall - bhd442"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
---

DONT FORGET GITHUB REPO

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = 'center')
```

```{r, include=FALSE}
library(tidyverse)
```

### Problem 1: Beauty, or not, in the classroom

```{r, include=FALSE}
profs <- read_csv("/Users/brandonduvall/Downloads/profs.csv")
```

#### Part A

```{r, fig.width=8, fig.height=6}
ggplot(profs) + geom_histogram(aes(x=eval), binwidth = 0.1) + labs(x="Course Eval Scores", y="Frequency", title="Distribution Of Course Eval Scores")
```

In this histogram, we can see that most students rated their provided instructors with a score of 4 or slightly below it, that is where the peak in the graph is. It is also somewhat skewed right with some scored being as low as a little below 2. Another key observation here is that no professor was ever rated with a course evaluation score of lower than 2 even though the rating range was 1 to 5. This could scale our findings as a 3 here in context may not be as high as it would be on a traditional 1 to 5 rating scale. We can takeaway here that many students are generous with their ratings in terms of course evaluations.

#### Part B

```{r, fig.width=8, fig.height=6}
ggplot(profs) + geom_boxplot(aes(x=native, y=eval)) + labs(x="Native English Speaker", y="Course Eval Scores", title="Distribution of Course Eval Scores By If Professor Is Native English Speaker")
```

The side-by-side boxplots show that on average, professors who spoke English as a native language had higher course evaluation scores by nearly half an additional score. The maxes of those who spoke English was also much higher than those who do not speak English, showing that the type of language spoken by professors could be important when observing their evaluation ratings. This also applies to how low some of their scores were, those who spoke English as a native language had some outstanding low points of 2.5 or less which was not directly attached to the boxplot. So, we can conclude that English speakers had a higher range of scores but on average had a higher score than those who did not speak English.

\newpage

#### Part C

```{r, fig.width=8,fig.height=6}
ggplot(profs) + geom_histogram(aes(x=eval), binwidth = 0.1) + facet_wrap(~gender, nrow=2) + labs(x="Course Eval Scores", y="Frequency", title="Distribution of Course Eval Scores by Gender")
```

The two histograms show that on average, the male professors had higher course evaluation scores but also lower minimums than female professors. The male average is likely around 4 while the female average is 3.75 or slightly more or less. Something else interesting to note is that not a single female professor got a rating of a 5 for their course evaluation score. With the two histograms, we are able to see that gender might have an effect on what students chose to rate their professors for their course evaluation scores.

\newpage

#### Part D

```{r}
ggplot(profs) + geom_point(aes(x=beauty,y=eval)) + labs(x="Physical Attractiveness",y="Course Eval Scores",title="Association Between Physical Attractives And Course Eval Scores")
```

For physical attractiveness, "2" indicates that the instructor is 2 points higher than the average beauty rating given by the panelists while "-1" would indicate 1 point lower than the average. The scatterplot here indicates that there is no strong but maybe a very weak correlation between a professor's physical attractiveness and their course evaluation scores as the dots are scattered completely. Professors who were great (had high course evaluation scores) were both "physically unattractive" (negative scores) and "physically attractive" (positive scores). The graph is able to support the idea that physical attractiveness does not connect to having a certain standard of course evaluation rating.

\newpage

### Problem 2: bike sharing

```{r, include=FALSE}
bikes <- read_csv("/Users/brandonduvall/Downloads/bikeshare.csv")
```

#### Plot A

```{r,fig.width=8,fig.height=6}
  bikes2 <- bikes %>%
  group_by(hr) %>%
  summarize(avg_bikehr = mean(total)) %>%
  round(0)
ggplot(bikes2) + geom_line(aes(x=hr,y=avg_bikehr)) + labs(x="Hours Of The Day",y="Average Count Of Bike Rentals That Hour",title="Change in Average Bike Rentals by Hours Of Day")
```

The line graph here follows the change in average bike rentals by hour for each day using the Capital Bikeshare system in Washington D.C. from the years 2011 and 2012. The x-axis shows the hours of the day with 1 being 1AM and 11 being 11AM and 13 being 1PM and so on. The y-axis is the average number of both casual and registered bike riders for each hour of the day. Looking at the graph, it is apparent that bike rental times are high in the mornings and also very high in the afternoon around when work ends. This would make sense because many citizens may use a bike as their transportation to and from work in a tight-knit city like Washington D.C. with a large public transportation network. Expectedly, bike rental averages are lowest very early in the morning and late at night because that is when many people are sleeping or not at work. Ridership is high during times when people need to get to and from work and low when most people are sleeping or not working.

\newpage

#### Plot B

```{r,fig.width=8,fig.height=6}
bikes3 <- bikes %>%
  group_by(hr, workingday) %>%
  summarize(avg_bikehr = mean(total), .groups = "drop") %>%
  round(0)
ggplot(bikes3) + geom_line(aes(x=hr,y=avg_bikehr)) + facet_wrap(~workingday) + labs(x="Hours Of The Day",y="Average Count Of Bike Rentals That Hour",title="Change in Average Bike Rentals by Hours Of Day")
```

This graph is very similar to the previous one as we are measuring the change in average ridership by hour. However, the main difference here is that we are doing this while also separating them on days that are weekdays (seen as 1) and days that are considered holidays/weekends (seen as 0). The point of making this facet is to see if common days of work affect the bike ridership. Looking at the graphs, it is clear to see that yes, there might be a difference as the weekdays graph has similar peaks as the previous graph from Plot A, while the holidays/weekends graph has higher ridership during the middle of day. It makes sense because people may be enjoying time off while not working during usual working hours. It can be concluded that the graph suggests that average ridership between weekdays and holidays/weekends differs likely due to work patterns.

\newpage

#### Plot C

```{r}
bikes4 <- bikes %>%
  filter(hr == 9) %>%
  group_by(weathersit, workingday) %>%
  summarize(avg_bikehr2 = mean(total), .groups = "drop") %>%
  round(0)
ggplot(bikes4) + geom_bar(aes(x=weathersit,y=avg_bikehr2), stat = "identity") + facet_wrap(~workingday) + labs(x="Weather Situation Code",y="Average Count of Bike Rentals At 9 AM",title="Average Ridership By Weather Code On Weekdays/Weekends During 9AM")
```

The two barplots here are meant to measure the average ridership for each weather situation code at 9 AM specifically as well as if there is a difference between the data on weekdays and weekends/holidays. To specify, 1 indicates: "Clear, Few clouds, Partly cloudy", 2 indicates: "Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist", 3 indicates: "Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds", 4 indicates: "Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog." Looking at the graphs, the average ridership was higher on weekdays for all weather situation codes during 9 AM meaning people are most likely still riding regardless of weather. However, a significantly smaller average is seen for weather situation codes at 3 because that is when the weather gets a little dangerous to ride a bike. Finally, the weather code was never 4 at 9 AM for any day. To conclude, ridership was smaller in general on weekends/holidays but the weather did seem to affect ridership for both groups in that the higher the code/worse the weather was, the ridership would decrease.

\newpage

### Problem 3: Capital Metro UT Ridership

```{r, include=FALSE}
capmetro_UT <- read_csv("/Users/brandonduvall/Downloads/capmetro_UT.csv")
capmetro_UT = mutate(capmetro_UT,
day_of_week = factor(day_of_week,
levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")),
month = factor(month,
levels=c("Sep", "Oct","Nov")))
```

#### Part 1

```{r, message=FALSE}
capmetro_UT2 <- capmetro_UT %>%
  group_by(hour_of_day, day_of_week, month) %>%
  summarize(avg_boarding = mean(boarding), .groups = "drop")
ggplot(capmetro_UT2) + geom_line(aes(x=hour_of_day,y=avg_boarding,color=month)) + facet_wrap(~day_of_week) + labs(x="Hours Of The Day", y="Average Boardings On UT's Capital Metro", title="Average Boardings on UT's Capital Metro By Hours and Month",color="Month")
```

Within the figure above, we are seeing the average number of boardings on the UT Capital Metro system per hour on each day of the week and how it changes throughout the day. Additionally, each singular graph has 3 colored lines to get the average boardings by hour of the day for the months of September, October, and November. Firstly, looking at the peaks, we can see that the peaks are pretty consistent on the weekdays at around hours 15-17 which makes sense as most people are using the capital metro system after their classes or working hours are over. The boardings reach about 150 on each of the weekdays on average for all months. However, on Saturdays and Sundays, there is no real peak but a consistent low number of boardings throughout the day (they reach about 50-60 a day on the weekends). Secondly, if we look at the number of boardings on Monday for the month of September, there is an obvious dip in average boardings for basically all hours of the day. This is most likely due to the Labor Day holiday that usually lands on the first Monday of September. As a result, there may be less bus drivers driving, the bus system may not be in operation that day, and students may not have to use the bus because of not having class that day. This would definitely bring the averages down for that line on that particular graph. Finally, when we look at the day Wednesday, Thursday, Friday for the month of November, we can see there the number of boardings are lower on average for all hours of the day. The cause of this is most likely Thanksgiving and Thanksgiving Break for students. As commonly known, Thanksgiving falls on a Thursday meaning that people the day before may be travelling to other places meaning the bus system is being used less and less. Additionally, the students of UT are on break meaning there are most likely less people in general using busses on those days.

\newpage

#### Part 2

```{r,fig.width=8,fig.height=6}
ggplot(capmetro_UT) + geom_point(aes(x=temperature,y=boarding, color = weekend),size=1) + facet_wrap(~hour_of_day) + labs(x="Temperature in Fahrenheit",y="Number Of People That Boarded UT's Capital Metro",title="Association Between Temperature And Number Of Boarders By Type Of Day",color="Day")
```

This graph is meant to show the potential links between temperature and the number of boardings related to the temperature by each hour of the day. There is also a color code for whether the days are weekend/holidays or are weekdays. Looking at all 16 graphs, the data is consistent with the line graphs as there are more data points around the middle of the afternoon. When keeping the hour of the day and type of day constant, we are able to observe that there is no relationship at all between the temperature in Fahrenheit and number of people boarding UT's Capital Metro System. For example, looking at just the hour 15 and weekdays (red dots), there is no relationship as the number of points does not change because of temperature, no correlation at all. Looking at all the other hours and their specific status, there is no connection either allowing us to conclude that there is no evidence of a relationship between the temperature in Fahrenheit and number of people that board UT's Capital Metro System.

\newpage

### Problem 4: Wrangling the Billboard Top 100

```{r, include=FALSE}
billboard <- read_csv("/Users/brandonduvall/Downloads/billboard.csv")
```

#### Part A

```{r}
billboard %>%
  group_by(performer, song) %>%
  summarize(count = n(), .groups = "drop") %>%
  arrange(desc(count)) %>%
  head(10)
```

THe table here shows the Top 10 songs that have spent the most weeks in the Billboard Top 100 from the years 1958 to 2021. As you can see, Radioactive by Imagine Dragons has spent the most weeks on the list at 87 weeks with the second highest being Sail by AWOLNATION with 79 weeks and so on and so forth. 

\newpage

#### Part B

```{r,fig.width=8,fig.height=6}
billboard2 <- billboard %>%
  filter(year != 1958 & year != 2021) %>%
  group_by(year, performer, song) %>%
  summarize(unique_by_sngprfrm = n(), .groups = "drop")
billboard3 <- billboard2 %>%
  group_by(year) %>%
  summarize(unique_by_year = n(), .groups = "drop")
ggplot(billboard3) + geom_line(aes(x=year,y=unique_by_year)) + labs(x="Year",y="Number Of Unique Songs On Billboard Top 100",title="Number Of Unique Songs On Billboard Top 100 By Year")
```

The next graph above analyses the number of unique songs that appeared on the Billboard Top 100 within that year. The change is measured in the form of a line graph. Looking at it, the unique number of songs increases from 1958 to 1967/68 where it reached its peak at a little over 850 unique songs. However, after this, it continues to decline until it reaches an all time low of 380/390 in the year 2001. This is very strange because it is less than half of its peak. After this low it rises almost back up to its peak at a little over 800 in the year 2020. The reason for the all time low is unknown but is interesting because it could have been caused by many things. Maybe there was a change in the style of music and popular music was transforming. Maybe there was less and less musicians during this time and unique songs weren't as frequent keeping the life of some songs longer than usual. Either way, it is cool to analyze how the number of unique songs in the Billboard Top 100 has changed over decades.

\newpage

#### Part C

```{r,fig.width=8,fig.height=6}
billboard4 <- billboard %>%
  group_by(performer, song) %>%
  summarize(count4 = n(), .groups = "drop") %>%
  arrange(desc(count4)) %>%
  filter(count4 >= 10) %>%
  group_by(performer) %>%
  summarize(ten_hits = n(), .groups = "drop") %>%
  filter(ten_hits >= 30)
ggplot(billboard4) + geom_bar(aes(x=performer,y=ten_hits),stat="identity") + labs(x="Name Of Performer",y="Number Of Ten Week Hits",title="Number Of Ten Week Hits Per Performer") + coord_flip()
```

Our final graph here is meant to view the 19 artists who have had 30 or more so called "ten-week hits" in the Billboard Top 100 from the years 1958 to 2021. First, we define a "ten-week hit" as a song that has spent at least 10 weeks or more on the Billboard Top 100 list. Each of these 19 artists has had at least 30 of these "ten-week hits," a monumental achievement that ranges all types of artists. Elton John tops the list with 54 "ten-week hits!" As said, there are modern artists on here like Drake and Taylor Swift. 