---
title: "SDS315HW7"
author: "Brandon Duvall"
date: "2025-04-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = 'center')
```

```{r, include=FALSE}
library(tidyverse)
library(mosaic)
library(MatchIt)
library(ggplot2)
```

[GitHub File Link](https://github.com/brandonduv/SDS315---Brandon-Duvall---HW/blob/main/SDS315HW7.Rmd)

### Problem 1: Armfolding

```{r, include=FALSE}
armfolds <- read_csv("/Users/brandonduvall/Downloads/armfold.csv")
```

#### Part A:

```{r, include=FALSE}
num_male <- armfolds %>%
  summarise(male_count = sum(Sex == "Male"))
num_male

num_female <- armfolds %>%
  summarise(female_count = sum(Sex == "Female"))
num_female

male_left <- armfolds %>%
  summarise(male_left_count = sum(Sex == "Male" & LonR_fold == "1"))

female_left <- armfolds %>%
  summarise(female_left_count = sum(Sex == "Female" & LonR_fold == "1"))

prop_male_left <- male_left/106
round(prop_male_left, 4)
prop_female_left <- female_left/111
round(prop_female_left, 4)
```

The number of male and female students in this dataset are 106 and 111 respectively.
The sample proportion of males who folded their left arm on top is 0.4717 while the sample proportion of females who folded their left arm on top is 0.4234

#### Part B:

```{r, include=FALSE}
diff_prop <- prop_male_left - prop_female_left
round(diff_prop, 4)
```

The observed difference in proportions between the two groups (males minus females) is 0.0483

#### Part C:

##### Built-in R

```{r}
left_CI = prop.test(LonR_fold ~ Sex, data=armfolds)
confint(left_CI)
```

This prints out the 95% confidence interval for the difference in proportions for males and females who folded their left hand on top using prop.test as -0.09 to 0.19.

##### Handwritten

```{r}
se_diffprop <- sqrt((0.4717*(1-0.4717)/106) + (0.4234*(1-0.4234)/111))
lower_bound <- 0.0483 - 1.96*(se_diffprop)
upper_bound <- 0.0483 + 1.96*(se_diffprop)
cat("Part 1 (Standard Error):", se_diffprop)
cat(" Part 2 (Lower Bound of CI):", lower_bound) 
cat(" Part 3 (Upper Bound of CI):", upper_bound)
```

Formula for Standard Error (SE) of Difference in Proportions: square root of (p1(1-p1)/n1 + p2(1-p2)/n2).

Here, p1 is the proportion of an occurence while 1-p1 gets the opposite proportion and n1 is the sample size of the first proportion. Then this is all added to the same thing for our second proportion function using p2, 1-p2, and n2. The square root of all of this is taken to find our SE (standard error). 

In this case, our 1 values are for the males meaning p1 = 0.4717, 1-p1 = 0.5283, and n1 = 106. Furthermore, our 2 values are for the females meaning p2 = 0.4234, 1-p2 = 0.5766, and n2 = 111. This gives us our standard error (SE(diffprop)) of 0.067. Finally, we plug this into our general confidence interval formula.

Formula for 95% Confidence Interval of Difference in Proportions: diffprop value +/- z-value * SE(diffprop)

For our problem, diffprop value = 0.0483, z-value = 1.96 (or 2), and SE(diffprop) = 0.067. We chose the z-value 1.96 because 95% (our confidence level) of the data lies within 1.96 standard deviations from our mean value (within a normal distribution).

Doing this gives us our confidence interval of -0.084 to 0.181 which is very similar to the result from the Built-in R function.

#### Part D:

If we were to run many many repeated random samples and calculated many differences in proportions, then we would expect that 95% of the confidence intervals contain the true difference in proportions of the population.

#### Part E:

The standard error we calculated above in this context represents how much our difference in proportions is expected to vary across many repeated samples drawn from the population. In this case, how much our difference in proportions of males and females who have their left arm on top fluctuates by from sample to sample by random variation (standard deviation of the sample).

#### Part F:

So the term sampling distribution here refers to the distribution of our sample difference in proportions of males to females who folded their left arm on top after running many samples. Since we are gathering samples, inherently the proportions of males and females who folded their left arm on top will differ meaning that the difference in proportion value in our sample will be different, creating a spread in our sampling distribution. The variance we see here is identified as the standard error, showing a range of difference of proportions we could expect to see in our sampling distribution. The things that stay fixed here are the sample size(s) of each sample we gather from our population of students at the Australian university.

#### Part G:

The mathematical result or theorem that justifies using a normal distribution to approximate the sampling distribution of the difference in sample proportions is called the Central Limit Theorem (CLT). In essense, the CLT states that sampling distributions of a sample statistic all end up being normal regardless of the population distribution as long as the sample size(s) are large enough. In this case, our sample sizes of 106 and 111 are more than sufficient to allow us to use a normal distribution which allows us to use a specific z-value to estimate our confidence intervals.

#### Part H:

To the person who says "there's no sex difference in arm folding," I would tend to support the statement because the difference in proportions of males and females who placed their left arm on top in a 95% confidence interval of -0.01 to 0.30 contains zero. This means that the interval is not statistically significant and supports the person's claim that there's no sex difference in arm folding. However, we could run more confidence intervals to be more certain about supporting this persons statement, however with just this interval, we can support their claim.

#### Part I:

If we were to repeat this experiment many times with different random samples of university students, the confidence interval would definitely vary and be different across our samples. The reason is that every time we get a new sample, we are drawing new males and females and as a result the proportions of those who fold their left arm on top between the two groups is going to be different, just due to random variation. This means our confidence interval for our sample difference in proportions will be different because our sample difference in proportions are different, even by small amounts. It should be true that if all the confidence intervals are done at 95%, that 95% of the confidence intervals for our difference in proportions of males and females who fold their left arm on top should cover the true difference in proportion value, also known as the coverage principle.

\newpage

### Problem 2: Get out the vote

```{r, include=FALSE}
turnout <- read_csv("/Users/brandonduvall/Downloads/turnout.csv")
```

#### Part A:

```{r}
prop(voted1998 ~ GOTV_call, data=turnout)
```

The proportion for those who voted in 1998 and did not receive a call was 0.444 while the proportion for those who also voted in 1998 but did receive a call was 0.647.

```{r}
GOTV_CI = prop.test(voted1998 ~ GOTV_call, data = turnout)
confint(GOTV_CI)
```

Our 95% confidence interval using prop.test for the difference in proportions for those who voted and did/did not receive a call is from 0.141 to 0.266. This is statistically significant because it does not include 0 and indicates a relationship between receiving/not receiving a GOTV_call and voting in the 1998 congressional election.

#### Part B:

```{r}
voted_summary = turnout %>%
  group_by(GOTV_call) %>%
  summarize(Vote1996 = mean(voted1996), Age = mean(AGE), Major = mean(MAJORPTY))
voted_summary
```

As seen in the table above, we can see a clear difference in proportions (mean for AGE) between the treatment variable (GOTV_call) and our other variables of interest. For whether or not receiving a GOTV_call correlated with whether or not someone voted in 1996, we can see that those who received a call and voted had a proportion of 0.71 while those who did not receive a call and voted had a proportion of 0.53. The same thing is seen for age with the mean age for those who got a call being 58.3 and 49.4 for those who did not. Finally, the proportion for those who were registered for a major party and did/did not receive a GOTV_call was 0.74 and 0.80. Next, we will examine each of these variables specifically and their associated confidence intervals/plots.

\newpage

##### Voted 1996 (Before Matching)

```{r}
VT1996_CI = prop.test(voted1996 ~ GOTV_call, data = turnout)
confint(VT1996_CI)
```

Our 95% confidence interval from 0.12 to 0.24 for the difference in proportions of voters in 1996 on whether or not they received a GOTV_call does not contain the value zero, meaning the interval is statistically significant and supports the claim of there being a correlation here.

\newpage

##### Age (Before Matching)

```{r, message=FALSE}
ggplot(turnout) +
  geom_histogram(aes(x=AGE,y=after_stat(density)),binwidth=2) +
  facet_wrap(~GOTV_call) + labs(x="Ages",y="Frequency (as proportion)",title="Age Distribution by GOTV call")
```

Looking at the distribution of ages in our graph, we can see that those who received a GOTV_call were definitely on average older than those who did not, providing support of age being a confounder because it correlates with whether or not someone received a GOTV_call.

```{r}
AGE_CI = t.test(AGE ~ GOTV_call, data = turnout)
confint(AGE_CI)
```

Additionally, our 95% confidence interval for the difference in mean ages between those who did not receive a GOTV_call and those who did is between -11.39 to -6.37 which is statistically significant since it does not include zero, further supporting the idea that age is a confounder in this case.

\newpage

##### Major Party (Before Matching)

```{r}
MJRPTY_CI = prop.test(MAJORPTY ~ GOTV_call, data = turnout)
confint(MJRPTY_CI)
```

The 95% confidence interval here for the difference in proportions of those who were registered for a major party and did/did not receive a GOTV_call is from 0.004 to 0.109. Since the interval does not include 0, we are statistically significant and able to support that there is likely a correlation between receiving a GOTV_call and being registered for a major party; being a potential confounding variable.

\newpage

#### Part C:

To really isolate whether or not receiving a GOTV_call had an impact on the proportion of those who voted in 1998, we can match our data for similarities in other variables.

```{r}
turnout_match = matchit(GOTV_call ~ voted1998 + voted1996 + AGE + MAJORPTY, data = turnout, ratio=5)
turnout_matched = match.data(turnout_match)
```

```{r}
voted_summary2 = turnout_matched %>%
  group_by(GOTV_call) %>%
  summarize(Vote1996 = mean(voted1996), Age = mean(AGE), Major = mean(MAJORPTY))
voted_summary2
```

After calculating a new table for our other variables and to see their potential differences in proportions (or mean) regarding whether or not they received a GOTV_call, we can now see that there is almost no difference. The difference of proportions of those who voted in 1996 and did/did not receive a GOTV_call goes to the third decimal point. The small difference is also seen in mean age (58.1 is very close to 58.3) and being registered for a major party (0.816 is very close to 0.802). This is evidence that they are no longer confounding variables in our matched data since there is no strong correlation.

\newpage

##### Voted 1996 (After Matching)

```{r}
match_VT1996_CI = prop.test(voted1996 ~ GOTV_call, data = turnout_matched)
confint(match_VT1996_CI)
```

Here, we can notice how the 95% confidence interval for the difference in proportions for those who voted in 1996 and did/did not receive a GOTV_call that was previously statistically significant is now no longer statistically significant since it ranges from -0.07 to 0.06 which includes 0, meaning it does not support the claim that voting in 1996 could be a confounding variable.

\newpage

##### Age (After Matching)

```{r}
ggplot(turnout_matched) +
  geom_histogram(aes(x=AGE,y=after_stat(density)),binwidth=2) +
  facet_wrap(~GOTV_call) + labs(x="Ages",y="Frequency (as proportion)",title="Age Distribution by GOTV call")
```
The faceted histogram using proportions (instead of count) also shows how similar our age variable and distribution are after matching, show essentially no difference between those who did/did not receive a GOTV_call.

```{r}
match_AGE_CI = t.test(AGE ~ GOTV_call, data = turnout_matched)
confint(match_AGE_CI)
```

Next, looking at our 95% confidence interval, we can see that the difference in mean ages of our matched data set ranges from -2.92 to 2.52 which includes 0. Since it includes 0, it is not statistically significant like it was before and supports the claim that it is no longer a confounding variable in our matched data.

\newpage

##### Major Party (After Matching)

```{r}
match_MJRPTY_CI = prop.test(MAJORPTY ~ GOTV_call, data = turnout_matched)
confint(match_MJRPTY_CI)
```

The 95% confidence interval for the difference in proportions of those who were registered for a major party and did/did not receive GOTV_call to illustrate this difference ranged from -0.07 to 0.04 meaning it is not statistically significant, further supporting the claim that GOTV_call is not correlated to being registered for a major party in our matched data set.

\newpage

##### Wrap-Up

```{r}
prop(voted1998 ~ GOTV_call, data=turnout_matched)
```

Going back to where this started using matched data set this time, we can see that there is a fourth decimal difference between the proportions of those who voted in 1998 and did not/did receive a GOTV_call.

```{r}
match_GOTV_CI = prop.test(voted1998 ~ GOTV_call, data = turnout_matched)
confint(match_GOTV_CI)
```

And with our 95% confidence interval, we can see that the difference in proportions for the voting in 1998 and whether or not someone received a GOTV_call ranges from -0.06 to 0.07 which includes 0. This means that this is not statistically significant and supports the idea that receiving a call had basically no impact on whether or not someone voted in 1998.

To wrap this up, we can quite confidently state that the GOTV_call had essentially no effect on the likelihood/proportion of voting in the 1998 election. We are able to conclude this after matching because the three other variables (voted1996, age, and major party) are no longer confounding variables allowing us to just see the relationship between GOTV_call and voted in 1998. And with the isolated data, we are able to conclude that there is no real relationship between whether or not someone received a GOTV_call and voting in 1998 using the confidence interval stated above.